---
title: "Philadelphia Legal Assistance: Adverse Possession"
author: "Adrian Leon, Max Masuda-Farkas, Gillian Xuezhu Zhao"
date: "28/01/2022"
output:
    html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---


# Philadelphia Legal Assistance

Smart Cities Practicum

## Introduction


```{r setup, include=FALSE}

# R Markdown options
knitr::opts_chunk$set(echo = T, warning = F, error = F, message = F, results = F, cache=T)

# LOAD PACKAGES

library(caret)
library(geojsonR)
library(gganimate)
library(gifski)
library(gridExtra)
library(kableExtra)
library(knitr)
library(lubridate)
library(mapview)
library(riem)
library(scales)
library(sf)
library(spdep)
library(tidycensus)
library(tidyverse)
library(tigris)
library(viridis)

# R options setup
options(scipen = 999)
options(tigris_class = "sf")

# additional functions from PPA book
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")


# function shortcuts
g <- glimpse
m <- mapview
len <- length
st_c <- st_coordinates


```




## Base data

### A. Geographic boundaries

```{r data: base geometry}

# set CRS to Pennsylvania South NAD83 in meters
phlcrs <- 'EPSG:32129'


# get geometry for tracts in Philadelphia
phltracts <- tigris::tracts(state = 42, county = 101) %>%
  dplyr::select(GEOID, geometry) %>%
  st_transform(st_crs(phlcrs))


# get geometry for the county of Philadelphia
phlcounty <- tigris::counties(state = 42) %>%
  filter(GEOID == '42101') %>%
  dplyr::select(GEOID, geometry) %>%
  st_transform(st_crs(phlcrs))


# set up the fishnet bins as the most granular unit of analysis
fishnet <- phlcounty %>%
  st_make_grid(.,
               cellsize = 250, 
               square = TRUE) %>%
  .[phlcounty] %>%                     # clip to Philadelphia County boundary
  st_sf() %>%
  mutate(uniqueID = rownames(.))


# Create a fishnet grid boundary of Philadelphia County
phlFishnet <- fishnet %>%
  st_union()

```





### B. Demographic

```{r data: ACS demographic}

# Get API key loaded locally
censusKey <-  read_file('../censusapikey.R') #"c2bdb8d298b60ec0cb50b561a2f8df61e92f6b39" 

# load census api key from local file
census_api_key(censusKey)

# set consulted year
year <- 2019

# get variables names and codes
acsVariableList <- load_variables(year, "acs5", cache = TRUE)

# set census variables to consult
censusVars <-
  c("B02001_001E", # Total population 
    "B11001_002E", # Total households
    "B02001_002E", # Total white population 
    "B19013_001E", # Median household income
    "B25003_001E") # Tenure of household

# Get the data
demographics <-
  get_acs(geography = "tract",
          variables = censusVars,
          year = year,
          state = 42,
          county = 101,
          geometry= T,
          output = "wide") %>%
  st_transform(st_crs(phlcrs)) %>%
  rename(totalPop = "B02001_001E",
         totalHHs = "B11001_002E",
         whitePop = "B02001_002E",
         medHHInc = "B19013_001E",
         tenureHH = "B25003_001E") %>%
  dplyr::select(-NAME, -starts_with("B"))%>%
  replace(is.na(.), 0) %>%
  mutate(pctWhite = ifelse(totalPop > 0, whitePop / totalPop, 0),
         pctTenur = ifelse(totalPop > 0, tenureHH / totalPop, 0)) %>%
  dplyr::select(-whitePop, -tenureHH)
  

# Create a mask for the majority minority census tracts fishnet cells
majmin <- 
  demographics %>%
  filter(pctWhite > 0.5) %>%
  st_transform(st_crs(phlcrs)) %>%
  st_union() %>%
  st_intersection(fishnet)

# Crop the fishnet by the majority minority census tracts
majminFishnet <- fishnet[majmin,] %>%
  st_union()


```



### C. Properties

[source]('https://www.opendataphilly.org/dataset/opa-property-assessments')


```{r data: OPA properties}

# Load properties from 99 until now from ODP.
# properties <- read_csv('https://opendata-downloads.s3.amazonaws.com/opa_properties_public.csv')


# Selected variables of interest
propertiesVars <-
  c(#"objectid",                       # UNNECESSARY
    #"assessment_date",                # UNNECESSARY
    #"basements",                      # UNNECESSARY
    #"beginning_point",                # UNNECESSARY
    #"book_and_page",                  # UNNECESSARY
    #"building_code",                  # UNNECESSARY
    #"building_code_description",      # UNNECESSARY
    "category_code",                  # KEEP ------------------------------------determines if it is VACANT LAND
    #"category_code_description",      # UNNECESSARY ----------------------------determines if it is VACANT LAND
    #"census_tract",                   # UNNECESSARY
    #"central_air",                    # UNNECESSARY
    #"cross_reference",                # UNNECESSARY
    #"date_exterior_condition",        # UNNECESSARY
    #"depth",                          # UNNECESSARY
    #"exempt_building",                # UNNECESSARY
    #"exempt_land",                    # UNNECESSARY
    "exterior_condition",             # KEEP ------------------------------------how the exterior appears based on observation.
    #"fireplaces",                     # UNNECESSARY
    #"frontage",                       # UNNECESSARY
    #"fuel",                           # UNNECESSARY
    #"garage_spaces",                  # UNNECESSARY
    #"garage_type",                    # UNNECESSARY
    #"general_construction",           # UNNECESSARY
    #"geographic_ward",                # UNNECESSARY
    #"homestead_exemption",            # UNNECESSARY
    #"house_extension",                # UNNECESSARY
    #"house_number",                   # UNNECESSARY
    #"interior_condition",             # UNNECESSARY
    "location",                       # KEEP for JOIN check****
    #"mailing_address_1",              # UNNECESSARY
    #"mailing_address_2",              # UNNECESSARY
    #"mailing_care_of",                # UNNECESSARY
    #"mailing_city_state",             # UNNECESSARY
    #"mailing_street",                 # UNNECESSARY
    #"mailing_zip",                    # UNNECESSARY
    "market_value",                   # KEEP ------------------------------------the certified market value of the property.
    #"market_value_date",              # UNNECESSARY
    #"number_of_bathrooms",            # UNNECESSARY
    #"number_of_bedrooms",             # UNNECESSARY
    #"number_of_rooms",                # UNNECESSARY
    #"number_stories",                 # UNNECESSARY
    #"off_street_open",                # UNNECESSARY
    #"other_building",                 # UNNECESSARY
    "owner_1",                        # KEEP?
    "owner_2",                        # KEEP?
    "parcel_number",                  # KEEP to JOIN***
    #"parcel_shape",                   # UNNECESSARY
    #"quality_grade",                  # UNNECESSARY
    #"recording_date",                 # KEEP? -----------------------------------date the deed was presented to records.
    "registry_number",                # KEEP to JOIN
    #"sale_date",                      # KEEP
    "sale_price",                     # KEEP
    #"separate_utilities",             # UNNECESSARY
    #"sewer",                          # UNNECESSARY
    #"site_type",                      # UNNECESSARY
    #"state_code",                     # UNNECESSARY
    #"street_code",                    # UNNECESSARY
    #"street_designation",             # UNNECESSARY
    #"street_direction",               # UNNECESSARY
    #"street_name",                    # UNNECESSARY
    #"suffix",                         # UNNECESSARY
    #"taxable_building",               # UNNECESSARY
    #"taxable_land",                   # UNNECESSARY
    #"topography",                     # UNNECESSARY
    "total_area",                     # KEEP
    #"total_livable_area",             # UNNECESSARY
    #"type_heater",                    # UNNECESSARY
    #"unfinished",                     # UNNECESSARY
    #"unit",                           # UNNECESSARY
    #"utility",                        # UNNECESSARY
    #"view_type",                      # UNNECESSARY
    "year_built",                     # KEEP
    #"year_built_estimate",            # UNNECESSARY
    #"zip_code",                       # UNNECESSARY
    "zoning",                         # KEEP
    "pin",                            # KEEP to JOIN ****
    "lat",                            # KEEP****
    "lng")                            # KEEP****

# Data wrangling
propertiesData <- properties %>%
  dplyr::select(propertiesVars) %>%
  rename("address" = location) %>%
  filter(!is.na(lat), !is.na(lng)) %>%              ### REMOVE before joining to parcels!!!
  st_as_sf(coords = c("lat","lng"), crs = 4326) %>%
  st_transform(st_crs(phlcrs))

```







## Supply-side data

### Vacant lots


From metadata:  "Input characteristics of the model are determined to be either positive or negative indicators of vacancy. A positive indicator has a tendency to signify vacancy (e.g., code violation issued by L+I). A negative indicator has a tendency to signify occupancy (e.g., OPA occupied property code). Individual indicators are weighted based on age of observation/record or other factors as determined by city department subject matter experts for a contributing dataset. Each indicator is also determined to represent a vacant lot or vacant building, as documented in the contributing dataset."

[source]('https://www.opendataphilly.org/dataset/vacant-property-indicators')


```{r supply vacant land}

# OpenDataPhilly Vacant Property Indicator - Lots
vacantLand <- read_csv('https://opendata.arcgis.com/datasets/19c35fb02d544a9bad0032b58268c9f9_0.csv')

# Select useful variables --- NO GEOMETRY
vacantLandVars <-
  c(#"OBJECTID",                       # UNNECESSARY
    "ADDRESS",                        # KEEP for JOIN CHECK
    #"OWNER1",                         # REPEATED (in Properties)
    #"OWNER2",                         # REPEATED (in Properties)
    "BLDG_DESC",                      # ???? ------------------------------------Building description from OPA
    "OPA_ID",                         # KEEP to JOIN****
    #"LNIADDRESSKEY",                  # UNNECESSARY-----------------------------Unique identifier from Licenses + Inspections
    #"COUNCILDISTRICT",                # UNNECESSARY
    "ZONINGBASEDISTRICT",             # KEEP***
    #"ZIPCODE",                        # UNNECESSARY
    "LAND_RANK")                      # KEEP?
    #"Shape__Area"                    # KEEP?
    #"Shape__Length")                  # UNNECESSARY

# Data wrangling
vacantLandData <- vacantLand %>%
  dplyr::select(vacantLandVars) %>%
  rename("address" = ADDRESS) %>%
  mutate(vacant = 'vacant')


# JOIN to propertiesData
# propertiesData <---> vacantLandData

len(unique(vacantLandData$parcel_number))

table(unique(vacantLandData$OPA_ID) %in% unique(propertiesData$parcel_number))
table(unique(propertiesData$parcel_number) %in% unique(vacantLandData$OPA_ID))

# 49/(27579+49)           # 0.2% of vacantLand rows lost
# 27579/(27579+553725)    # 4.7% of properties joined to vacantLand

vacantLandProps <- propertiesData %>%
  inner_join(vacantLandData, by=c('parcel_number'='OPA_ID'))

# save locally
# saveRDS(vacantLandProps, file = "./local/vacantLandProps.rds")

```




### Real Estate Tax Delinquencies

[source]('https://www.opendataphilly.org/dataset/property-tax-delinquencies')


```{r supply tax delinquencies}

# Real Estate Delinquencies from OpenData Philly:
delinquencies <- read_csv('https://phl.carto.com/api/v2/sql?q=SELECT+*+FROM+real_estate_tax_delinquencies&filename=real_estate_tax_delinquencies&format=csv&skipfields=cartodb_id,the_geom,the_geom_webmercator')

# Variables - Tax Delinquencies (53 vars)
delinquenciesVars <-
  c(#"objectid",                       # UNNECESSARY
    "opa_number",                     # KEEP to JOIN****
    "street_address",                 # KEEP to JOIN check****
    #"zip_code",                       # UNNECESSARY
    #"zip_4",                          # UNNECESSARY
    #"unit_type",                      # UNNECESSARY
    #"unit_num",                       # UNNECESSARY
    #"owner",                          # UNNECESSARY
    #"co_owner",                       # UNNECESSARY
    #"principal_due",                  # UNNECESSARY -----------------------------The amount of principal due.
    #"penalty_due",                    # UNNECESSARY -----------------------------The penalty amount due.
    #"interest_due",                   # UNNECESSARY -----------------------------The amount of interest due.
    #"other_charges_due",              # UNNECESSARY -----------------------------Additional charges owner owes.
    "total_due",                      # KEEP ------------------------------------The total amount owner owes.
    "is_actionable",                  # KEEP?*** --------------------------------Action can be taken against the delinquent property.
    #"payment_agreement",              # UNNECESSARY -----------------------------The property owner is in a payment agreement.
    "num_years_owed",                 # KEEP ***for adverse possession reference
    "most_recent_year_owed",          # KEEP for reference
    "oldest_year_owed",               # KEEP for reference
    "most_recent_payment_date",       # KEEP for reference
    "year_of_last_assessment",        # KEEP for reference
    "total_assessment",               # KEEP for reference
    #"taxable_assessment",             # UNNECESSARY
    #"mailing_address",                # UNNECESSARY
    #"mailing_city",                   # UNNECESSARY
    #"mailing_state",                  # UNNECESSARY
    #"mailing_zip",                    # UNNECESSARY
    #"return_mail",                    # UNNECESSARY
    #"building_code",                  # ???? ------------------------------------Building codes describe the building.
    #"detail_building_description",    # ???? ------------------------------------Description of the building code.
    #"general_building_description",   # ???? ------------------------------------General description of how the building is used.
    "building_category",              # ???? ------------------------------------Type of Building Group (residential, commercial, etc.).
    #"coll_agency_num_years",          # ???? ------------------------------------TO COLLECTION AGENCY
    #"coll_agency_most_recent_year",   # ???? ------------------------------------TO COLLECTION AGENCY
    #"coll_agency_oldest_year",        # ???? ------------------------------------TO COLLECTION AGENCY
    #"coll_agency_principal_owed",     # ???? ------------------------------------TO COLLECTION AGENCY
    #"coll_agency_total_owed",         # ???? ------------------------------------TO COLLECTION AGENCY
    #"exempt_abatement_assessment",    # UNNECESSARY -----------------------------The amount excluded from the tax base.
    #"homestead_value",                # UNNECESSARY -----------------------------Amount of Homestead Exemption.
    #"net_tax_value_after_homestead",  # ???? ------------------------------------The net tax value after the homestead exemption.
    #"agreement_agency",               # UNNECESSARY -----------------------------Either "TIPS","lbr","grb","LAW" 
    #"sequestration_enforcement",      # UNNECESSARY -----------------------------If the property is a rental, the city can collect rent
    #"bankruptcy",                     # UNNECESSARY -----------------------------Active bankruptcy
    #"years_in_bankruptcy",            # UNNECESSARY 
    #"most_recent_bankrupt_year",      # UNNECESSARY
    #"oldest_bankrupt_year",           # UNNECESSARY
    #"principal_sum_bankrupt_years",   # UNNECESSARY
    #"total_amount_bankrupt_years",    # UNNECESSARY
    "sheriff_sale",                   # KEEP*** ---------------------------------Property is in the Sheriff Sale Process (any stage).
    "liens_sold_1990s",               # KEEP*** ---------------------------------Property was included in 1997 Lien Sale.
    "liens_sold_2015")                # KEEP*** ---------------------------------Property is included in Recent Lien Sales.
    #"assessment_under_appeal",        # UNNECESSARY -----------------------------the owner has appealed the assessment.
    #"year_month")                     # year$month >>>> Unlisted in metadata


# Data wrangling
delinquenciesData <- delinquencies %>%
  dplyr::select(delinquenciesVars) %>%
  rename("address" = street_address) %>%
  mutate(opa_number = sprintf("%09d", opa_number))


# JOIN to properties
# propertiesData <---> delinquenciesData

len(unique(delinquenciesData$opa_number))

table(unique(delinquenciesData$opa_number) %in% unique(propertiesData$parcel_number))
table(unique(propertiesData$parcel_number) %in% unique(delinquenciesData$opa_number))

# 2077/(70639+2077)       #  3%
# 70639/(70639+510665)    # 12%

initialDate = '2020-01-01'
finalDate = '2021-01-01'

delinquenciesProps <- propertiesData %>%
  inner_join(delinquenciesData, by = c('parcel_number'='opa_number')) %>%
  filter(sale_date >= initialDate & sale_date < finalDate) %>%
  st_sf()

# save locally
# saveRDS(delinquenciesProps, file = "./local/delinquenciesProps.rds")

```






### US Bank Liens

```{r data: us bank liens}

# Client-provided US Bank Liens Data (as 'US Bank sharp')
# These are Liens that have been sold - UPDATE?
usBankLiens_0 <- read.csv("./data/usBank.csv") %>%
  dplyr::select(-X) %>%
  na.omit() %>%
  st_as_sf(coords = c('x', 'y'), crs = 4326, agr = 'constant')


# US Bank Liens geocoded (by GILLIAN):
usBankLiens_1 <- read.csv("./data/usBankLiens_latlon.csv") %>%
  rename(lon = 'long') %>%
  filter(!is.na(lat), !is.na(lon)) %>% 
  st_as_sf(coords = c("lon","lat"), crs = 4269)  %>%
  filter(X != 16)                                                # TAKE out Mississippi geocode error


# US Bank Liens pending geocoding (with NAs):
usBankLiens_NA <- read.csv("./data/usBankLiens_latlonNA.csv") %>%
  rename(lon = 'long')


# visualize liens locations
m(usBankLiens_1, zcol='CALC_TOTAL')


```







## Demand-side data

### L&I Building and Zoning Permits

[source]('https://www.opendataphilly.org/dataset/licenses-and-inspections-building-permits')

```{r data: permits}

# DEMAND SIDE DATA
initialDate <- as.Date.character('2017-01-01')

# Load PERMITS data from OpenDataPhilly's Carto API
permits <- read_csv('https://phl.carto.com/api/v2/sql?q=SELECT+*,+ST_Y(the_geom)+AS+lat,+ST_X(the_geom)+AS+lng+FROM+permits&filename=permits&format=csv&skipfields=cartodb_id')

# List the permit data set variables needed
varsPermits <-
  c(#'objectid',                      # UNNECESSARY
    #'permitnumber',                  # UNNECESSARY
    #'addressobjectid',               # UNNECESSARY
    'parcel_id_num',                 # KEEP for JOIN
    'permittype',                    # KEEP*****
    'permitdescription',             # KEEP for reference
    'commercialorresidential',        # MAYBE
    'typeofwork',                    # KEEP****
    #'approvedscopeofwork',           # UNNECESSARY - detailed description
    'permitissuedate',               # KEEP***
    'status',                        # KEEP
    'applicanttype',                 # MAYBE
    #'contractor[...]',               # UNNECESSARY
    #'mostrecentinsp',                # UNNECESSARY
    'opa_account_num',               # KEEP for JOIN****
    'address',                       # KEEP for JOIN check
    #'unit_type',                     # UNNECESSARY
    #'unit_num',                      # UNNECESSARY
    #'zip',                           # UNNECESSARY
    #'censustract',                   # UNNECESSARY
    #'council_district',              # UNNECESSARY
    #'opa_owner',                     # REPETITIVE ------------------------------OPA's ownership from the current deed for the property.
    #'systemofrecord',                # UNNECESSARY
    #'posse_jobid',                   # UNNECESSARY
    'lng','lat'
    #'geometry'
    )

# EXAMINE VARIABLES
# permittype
permitTypes <- as.data.frame(table(permits$permittype))

# select permittyoe
# All categories
permitCats <- 
  c(#'ADMINISTRATIVE',        #    2030
    #'BP_ADDITON',            #    1781
    #'BP_ADMINST',            #    6499
    #'BP_ALTER',              #   33475
    #'BP_DEMO',               #    6259
    #'BP_FIRESUP',            #    9553
    #'BP_MECH',               #   18747
    'BP_NEWCNST',             #    4984   # "2007-01-02" to "2020-02-20"
    #'BP_SIGN',               #    1546
    'BUILDING',               #    8521   # "2015-01-12" to NOW ("2022-01-29")
    #'DEMOLITION',            #    1704
    #'ELECTRICAL',            #   22321
    #'EP_ELECTRL',            #   36433
    #'FIRE SUPPRESSION',      #    7089
    #'GENERAL',               #    1472 
    #'GENERAL PERMIT MINOR',  #    9225
    #'MASTER PLAN',           #      29
    #'MECHANICAL',            #   12662
    #'OPERATIONS',            #     579
    #'OPS PERMIT',            #    2232
    #'PLUMBING',              #   31451
    #'PP_PLUMBNG',            #   52541
    #'PRELIMINARY APPROVAL',  #      27
    'RESIDENTIAL BUILDING',   #   15909   # "2015-01-06" to NOW ("2022-01-29")
    #'SITE / UTILITY PERMIT', #     415
    'ZONING',                 #   11187   # "2015-12-02" to NOW ("2022-01-29")
    'ZP_ADMIN',              #     803
    'ZP_USE',                 #    9651   # "2007-01-02" to "2020-03-12"
    'ZP_ZON/USE',             #   12921   # "2007-01-02" to "2020-03-12"
    'ZP_ZONING'               #    7617   # "2007-01-02" to "2020-03-12"
    )

# Old categories
permitCatsA <- 
  c('BP_NEWCNST',             #    4984   # "2007-01-02" to "2020-02-20"
    'ZP_ADMIN',               #     803
    'ZP_USE',                 #    9651   # "2007-01-02" to "2020-03-12"
    'ZP_ZON/USE',             #   12921   # "2007-01-02" to "2020-03-12"
    'ZP_ZONING')              #    7617   # "2007-01-02" to "2020-03-12"

# New categories
permitCatsB <- 
  c('BUILDING',               #    8521   # "2015-01-12" to NOW ("2022-01-29")
    'RESIDENTIAL BUILDING',   #   15909   # "2015-01-06" to NOW ("2022-01-29")
    'ZONING')                 #   11187   # "2015-12-02" to NOW ("2022-01-29")


# Building categories
permitBuildingCats <- 
  c('BP_NEWCNST',             #    4984   # "2007-01-02" to "2020-02-20"
    'BUILDING',               #    8521   # "2015-01-12" to NOW ("2022-01-29")
    'RESIDENTIAL BUILDING')   #   15909   # "2015-01-06" to NOW ("2022-01-29")

# Zoning categories
permitZoningCats <- 
  c('ZP_ADMIN',               #     803
    'ZP_USE',                 #    9651   # "2007-01-02" to "2020-03-12"
    'ZP_ZON/USE',             #   12921   # "2007-01-02" to "2020-03-12"
    'ZP_ZONING',              #    7617   # "2007-01-02" to "2020-03-12"
    'ZONING')                 #   11187   # "2015-12-02" to NOW ("2022-01-29")


# Data wrangling ONE
# Filter data with PERMIT TYPES (building or zoning)
permitsData <- permits %>%
  dplyr::select(varsPermits) %>%
  filter(permittype %in% permitZoningCats) %>%
  filter(!is.na(lat), !is.na(lng)) %>% 
  st_as_sf(coords = c("lng","lat"), crs = phlcrs) 

# type of work across SELECTED categories (48):
permitTypesOfWork <- as.data.frame(table(permitsData$typeofwork)) 

# All categories
permitTypeOfWorkCats <- 
  c(#'ADD',                                              #   1665 ***OLD ***ZONING {New construction attached or added to existing}
    #'ADDITION AND/OR ALTERATION',                       #  17051 ***NEW ***BUILDING (6290)  ***RESIDENTIAL BUILDING (10761)
    #'ALTER',                                            #      1 ***OLD ***ZONING
    #'AMEND',                                            #    162 ***OLD ***ZONING {Admin stuff}
    #'CHANGE OF USE',                                    #   3046 ***NEW ***ZONING {General change of use}
    'COMBINED LOT LINE RELOCATION AND NEW DEVELOPMENT', #    313 ***NEW ***ZONING {Combine or redraw lots} ***SIGNIFICATIVE!
    'COMDEM',                                           #   1163 ***OLD ***ZONING {Complete demolition}
    #'CONDITIONAL ZONING APPROVAL',                      #     30 ***NEW ***ZONING
    #'CONR',                                             #      1 ***OLD ***ZONING
    #'COVID-19 TEMP USE (RESTAURANTS / RETAIL)',         #     71 ***NEW ***ZONING
    #'ENTIR',                                            #      3 ***OLD ***ZONING (-1)    ***BUILDING
    'ENTIRE',                                           #   2910 ***OLD ***ZONING (-2890) ***BUILDING {Entire structure}
    'ENTSTR',                                           #   3638 ***OLD ***ZONING
    #'FAMDAY',                                           #    268 ***OLD ***ZONING
    #'FAMILY DAYCARE',                                   #     65 ***NEW ***ZONING
    #'FENCE',                                            #     40 ***OLD ***ZONING
    #'FENCES ONLY',                                      #     26 ***NEW ***ZONING
    #'FITOUT',                                           #     91 ***OLD ***BUILDING {Commercial interiors}
    #'FOUND',                                            #    168 ***OLD ***BUILDING {Just foundations}
    #'FRAME',                                            #      1 ***OLD ***BUILDING
    'FULL DEMOLITION',                                  #    644 ***NEW ***ZONING {Full demolition}
    'LOT LINE RELOCATION',                              #    317 ***NEW ***ZONING {Combine or redraw lots} ***SIGNIFICATIVE!
    'LOTLIN',                                           #   2527 ***OLD ***ZONING {Combine or redraw lots} ***SIGNIFICATIVE!
    #'MAJOR',                                            #      4 ***OLD ***ZONING (-3) ***BUILDING
    #'MOD/IN',                                           #     17 ***OLD ***BUILDING {Modular structure}
    'NEW CONSTRUCTION',                                 #<- 6292 ***NEW ***BUILDING (1812) ***RESIDENTIAL BUILDING (4480)
    'NEW CONSTRUCTION (SHELL ONLY)',                    #<-   19 ***NEW ***BUILDING
    #'NEW CONSTRUCTION (STAND ALONE)',                   #<-   48 ***NEW ***RESIDENTIAL BUILDING
    'NEW CONSTRUCTION, ADDITION, GFA CHANGE',           #<- 4986 ***NEW ***ZONING {New construction attached or added to existing}
    'NEWCON'#                                           #<- 5539 ***OLD ***ZONING (-3) ***BUILDING {New construction of structure}
    #'PARKING ONLY',                                     #<-   97 ***NEW ***ZONING
    #'PARTBL',                                           #    116 ***OLD ***ZONING {Partial demolitions}
    #'PARTCH',                                           #   3317 ***OLD ***ZONING {Change in use}
    #'PARTIA',                                           #      4 ***OLD ***ZONING {Partial or complete demolitions}
    #'PARTSI',                                           #     36 ***OLD ***ZONING {Partial or complete demolitions}
    #'PRELIM',                                           #      6 ***OLD ***ZONING {Preliminary study}
    #'PREPL',                                            #      1
    #'SFADD',                                            #   3149 ***OLD ***ZONING {Add square footage to existing}
    #'SFRED',                                            #    253 ***OLD ***ZONING {Reduce square footage to existing}
    #'SHELL',                                            #      9 ***OLD ***BUILDING
    #'SIGNAC',                                           #   1584 ***OLD ***ZONING
    #'SIGNE',                                            #     19 ***OLD ***ZONING
    #'SIGNNA',                                           #     18 ***OLD ***ZONING
    #'SIGNS (ACCESSORY / NON-ACCESSORY)',                #    872 ***NEW ***ZONING
    #'SITE',                                             #     29 ***OLD ***BUILDING
    #'TANK',                                             #     27 ***OLD ***ZONING (-26) ***BUILDING
    #'TANKRI',                                           #      1 ***OLD ***BUILDING
    #'ZBAADM'                                            #    332 ***OLD ***ZONING {Admin stuff}
    )
    

# Data wrangling TWO
# Filter data with PERMIT TYPES (building or zoning)
permitsZoningData <- permitsData %>%
  filter(typeofwork %in% permitTypeOfWorkCats)


# JOIN to properties
# permitsZoningData <---> propertiesData

len(unique(permitsZoningData$opa_account_num))

table(unique(permitsZoningData$opa_account_num) %in% unique(propertiesData$parcel_number))
table(unique(propertiesData$parcel_number) %in% unique(permitsZoningData$opa_account_num)) 

# 540/(31466+540)         # 1.7% of permits rows lost
# 31466/(31466+549838)    # 5.4% of properties joined to permits

permitsProps <- propertiesData %>%
  inner_join(st_drop_geometry(permitsZoningData),  by = c('parcel_number'='opa_account_num'))

# permits by quarter by fishnet cell
permitsNet <- permitsProps %>%
  st_join(fishnet, left=F) %>%
  mutate(year = year(permitissuedate),
         month = month(permitissuedate),
         quarter = case_when(
           month %in% c(1, 2, 3) ~ 1,
           month %in% c(4, 5, 6) ~ 2,
           month %in% c(7, 8, 9) ~ 3,
           month %in% c(10, 11, 12) ~ 4)) %>%
  mutate(period = paste(year, quarter, sep='-')) %>%
  group_by(period, uniqueID) %>%
  summarize(permitCount = sum(n())) %>%
  st_drop_geometry()

# random cross validation ID (out of 100) for later use
# mutate(uniqueID = rownames(.)) %>%
# mutate(cvID = sample(round(nrow(fishnet)/24), size=nrow(fishnet), replace=T)) 

# create empty panel with all possible time/space combinations by fishnet grids
basePanel <- 
  expand.grid(period = unique(permitsNet$period), 
              uniqueID = unique(fishnet$uniqueID))

# join panel to fishnet geometry
permitsPanel <- 
  permitsNet %>%
  right_join(basePanel) %>% 
  replace(is.na(.), 0) %>%
  left_join(fishnet) %>%
  st_sf()

```




### Permits in the last 15 years

```{r data: permit issue evolution A, fig.width= 8, fig.height= 8}

# Create Animation
animation <- 
  ggplot() +
  geom_sf(data = permitsPanel, aes(fill = permitCount), color = NA) +
  scale_fill_viridis(option = "magma",
                     limits = c(0, 20),
                     trans = "log1p",
                     breaks = c(0, 5, 10, 20),
                     direction = 1) +
  labs(title = "Zoning permits issued in Philadelphia",
       fill = "",
       caption = "Source: OpenDataPhilly") +
  guides(fill= guide_colorbar(barheight = 9, title.position = "bottom", label.position = "right")) +
  geom_text(data = permitsPanel,
            aes(label = substr(period, 1, 4)),
            size = 8, x = -Inf, y = Inf, vjust = 1.5, hjust = -.1, colour = "black") +
    transition_manual(period) +
    mapTheme() +
    theme(legend.position = c(0.85, 0.2),
          legend.background = element_blank(),
          panel.border = element_blank(),
          panel.background = element_rect(fill = "#eeeeee"),
          panel.grid = element_blank()
          )

animation

# anim_save("testPLA.gif", animation)

```




```{r data: permits issue change: 5 years A, fig.width= 12, fig.height= 8}

# classify zoning permits by five-year periods
permits5years <- 
  permitsPanel %>%
  mutate(year = as.integer(substr(period, 1,4)),
         lustrum = case_when(
           year %in% seq(2007, 2011) ~ '2007-2011',
           year %in% seq(2012, 2016) ~ '2012-2016',
           year %in% seq(2017, 2021) ~ '2017-2021'
         )) %>%
  filter(year != '2022') %>%
  dplyr::select(-period, -year) %>%
  group_by(lustrum, uniqueID) %>%
  summarize(permitCount = sum(permitCount))


# side-by-side five-year period zoning permits
permits5years %>%
  ggplot() +
  geom_sf(aes(fill = permitCount), color = NA) +
  scale_fill_viridis(
    option = "magma",
    limits = c(0,120),
    trans = "log1p",
    breaks = c(0,30,60,120),
    direction = 1) +
  labs(
    title = "Zoning permits issued in Philadelphia",
    fill = "number of permits",
    caption = "Source: OpenDataPhilly") +
  guides(
    size = F,
    fill= guide_colorbar(barwidth = 21, title.position = "top")) +
  facet_wrap(~lustrum, ncol = 3) +
  mapTheme() +
  theme(
    legend.position = "bottom",
    legend.background = element_blank(),
    panel.border = element_blank(),
    panel.background = element_rect(fill = "#eeeeee"),
    panel.grid = element_blank(),
    strip.background = element_rect(fill = "#eeeeee"),
    strip.text.x = element_text(size = 12, color = '#222222', hjust=0.01)
    )

```



```{r data: permits issue change: 5 years B}

plotPermitsMap <- function(panel, start, fillcolor) {
  panel %>%
  mutate(year = as.numeric(substr(period, 1,4))) %>%
  filter(between(year, start, start + 4)) %>%
  group_by(uniqueID) %>%
  summarize(permitCount = sum(permitCount)) %>%
  ggplot() +
  geom_sf(aes(fill = permitCount), color = NA) +
  scale_fill_viridis(option = "magma",
                       limits = c(0,120),
                       trans = "log1p",
                       breaks = c(0,30,60,120),
                       direction = 1) +
  geom_sf(data = majminFishnet, fill = fillcolor, color = NA, alpha = 1) +
  geom_sf(data = phlFishnet, fill = NA, colour = "#222222") +
  labs(title = "Zoning permits issued in Philadelphia",
       subtitle = paste0("Period: ", start, '-', (start+4)),
       fill = "",
       caption = "Source: OpenDataPhilly") +
  guides(size = F,
         fill= guide_colorbar(barheight = 13, title.position = "bottom", label.position = "right")) +
  mapTheme() +
  theme(legend.position = c(0.85, 0.2),
        legend.title = element_text(size = 10, color = '#222222', hjust = 0.5),
        legend.background = element_blank(),
        legend.text.align = 0.9,
        legend.text = element_text(size=8),
        panel.border = element_blank(),
        panel.background = element_rect(fill = "#eeeeee"),
        panel.grid = element_blank()
        ) 
}

```


```{r fig.width= 8, fig.height= 8}

# permits in 5 year intervals
plotPermitsMap(permitsPanel, 2007, NA)
plotPermitsMap(permitsPanel, 2012, NA)
plotPermitsMap(permitsPanel, 2017, NA)

```


```{r fig.width= 8, fig.height= 8}

# permits in 5-year intervals with white majority masked out
plotPermitsMap(permitsPanel, 2007, "#eeeeee")
plotPermitsMap(permitsPanel, 2012, "#eeeeee")
plotPermitsMap(permitsPanel, 2017, "#eeeeee")

```





### Real Estate Transfers

Real Estate Transfers Tax Documents

[source]('https://www.opendataphilly.org/dataset/real-estate-transfers')

Dataset description:

"Description	This is a summary of real estate transfer tax (RTT) documents. This table contains both raw source data as well as calculated/geocoded data, which are described below. Each row represents a distinct property involved in a RTT transaction document; all properties involved in a transaction share the same document_id. Please note that 918 transactions have been temporarily omitted while we address data integration issues with these records."


```{r data: transfers}

# OpenDataPhilly Endpoint of Property Assessment History:
#transfers <- read_csv('https://opendata-downloads.s3.amazonaws.com/rtt_summary.csv')
transfers <- read_csv('https://opendata.arcgis.com/datasets/88e5bc291b834606bd49f6fd6dca226e_0.csv')

# OpenDataPhilly - Transfers (48 variables)
transfersVars <- 
  c(#"objectid",                     # UNNECESSARY
    #"document_id",                  # UNNECESSARY -------------------------------id of the real estate transfer document 
    "document_type",                # KEEP*** -----------------------------------refers to type of Real Estate Transaction
    "display_date",                 # KEEP***
    "street_address",               # KEEP for JOIN check
    #"zip_code",                     # UNNECESSARY
    #"ward",                         # UNNECESSARY
    "grantors",                     # KEEP? -------------------------------------seller (on deeds), or borrower (on mortgages)
    "grantees",                     # KEEP? -------------------------------------buyer, recipient, new owner, or lien holder
    #"cash_consideration",           # UNNECESSARY
    #"other_consideration",          # UNNECESSARY
    #"total_consideration",          # KEEP?** -----------------------------------good exchanged for the real estate (usually money)
    #"assessed_value",               # KEEP?** -----------------------------------assess value by OPA
    #"common_level_ratio",           # UNNECESSARY -------------------------------fair market : assessment values ratio
    "fair_market_value",            # KEEP?** -----------------------------------assessment value by common level ratio
    #"state_tax_amount",             # UNNECESSARY
    #"state_tax_percent",            # UNNECESSARY
    #"local_tax_amount",             # UNNECESSARY
    #"local_tax_percent",            # UNNECESSARY
    #"adjusted_cash_consideration",  # UNNECESSARY
    #"adjusted_other_consideration", # UNNECESSARY
    #"adjusted_total_consideration", # UNNECESSARY
    #"adjusted_assessed_value",      # UNNECESSARY
    #"adjusted_fair_market_value",   # UNNECESSARY
    #"adjusted_state_tax_amount",    # UNNECESSARY
    #"adjusted_local_tax_amount",    # UNNECESSARY
    #"receipt_num",                  # UNNECESSARY
    #"receipt_date",                 # UNNECESSARY - display_date is used
    #"recording_date",               # UNNECESSARY - display_date is used
    #"document_date",                # UNNECESSARY - display_date is used
    #"condo_name",                   # UNNECESSARY
    #"unit_num",                     # UNNECESSARY
    #"address_low",                  # UNNECESSARY
    #"address_low_suffix",           # UNNECESSARY
    #"address_low_frac",             # UNNECESSARY
    #"address_high",                 # UNNECESSARY
    #"street_predir",                # UNNECESSARY
    #"street_name",                  # UNNECESSARY
    #"street_suffix",                # UNNECESSARY
    #"street_postdir",               # UNNECESSARY
    #"reg_map_id",                   # UNNECESSARY ------------------------------DOR parcel ID sourced from CRIS_Properties.
    #"matched_regmap",               # UNNECESSARY ------------------------------DOR parcel ID matched to the address by the geocoder.
    "opa_account_num",              # KEEP****to join
    #"legal_remarks",                # UNNECESSARY
    #"discrepancy",                  # UNNECESSARY
    "property_count")               # KEEP --------------------------------------Number of properties in document
    #"lat",                          # KEEP
    #"lng")                          # KEEP

# transfersData <- transfers %>%
#   filter(display_date > initialDate) %>%
#   dplyr::select(transfersVars) %>%
#   rename('address' = street_address) %>%
#   filter(!is.na(opa_account_num)) %>%
#   filter(document_type %in% c('DEED',
#                               'DEED LAND BANK',
#                               'DEED MISCELLANEOUS',
#                               'DEED MISCELLANEOUS TAXABLE',
#                               'DEED OF CONDEMNATION',
#                               'DEED SHERIFF'))

# Data wrangling
transfersData <- transfers %>%
  filter(DISPLAY_DATE > initialDate) %>%
  dplyr::select(toupper(transfersVars)) %>%
  rename('address' = STREET_ADDRESS) %>%
  filter(!is.na(OPA_ACCOUNT_NUM)) %>%
  filter(DOCUMENT_TYPE %in% c('DEED',
                              'DEED LAND BANK',
                              'DEED MISCELLANEOUS',
                              'DEED MISCELLANEOUS TAXABLE',
                              'DEED OF CONDEMNATION',
                              'DEED SHERIFF'))


# JOIN to properties
# propertiesData <---> transfersData

len(unique(transfersData$opa_account_num))

table(unique(transfersData$opa_account_num) %in% unique(propertiesData$parcel_number))
table(unique(propertiesData$parcel_number) %in% unique(transfersData$opa_account_num))

424/(58427+425)       #  1%
58427/(58427+522877)  # 10%

transfersProps <- propertiesData %>%
  inner_join(transfersData, by= c('parcel_number'='opa_account_num'))
  
# save locally
#csaveRDS(transfersProps, file = "./local/transfersProps.rds")

```




### Transit Data

```{r data: transit access}

##  "Distance to transit?"

# Load SEPTA stops data
# get subway stops
subwayStops <- read.csv('./data/SEPTA_-_Highspeed_Stations.csv') %>%
  st_as_sf(coords = c("Longitude","Latitude"), crs = 4269) %>%
  st_transform(st_crs(phlcrs)) %>%
  st_filter(., phlcounty) %>%
  mutate(substops = 1) %>% 
  dplyr::select(substops, geometry)
  
# get trolley stops
trolleyStops <- read.csv('./data/SEPTA_-_Trolley_Stops.csv') %>%
  st_as_sf(coords = c("Longitude","Latitude"), crs = 4269) %>%
  st_transform(st_crs(phlcrs)) %>%
  st_filter(., phlcounty) %>%
  mutate(trostops = 1) %>% 
  dplyr::select(trostops, geometry)


```










