---
title: "Philadelphia Legal Assistance: Adverse Possession"
author: "Adrian Leon, Max Masuda-Farkas, Gillian Xuezhu Zhao"
date: "28/01/2022"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---


# Philadelphia Legal Assistance

Smart Cities Practicum



```{r setup, include=FALSE}

# R Markdown options
knitr::opts_chunk$set(echo = T, warning = F, error = F, message = F, results = F, cache=T, cache.lazy = F)

# LOAD PACKAGES
library(basemaps)
library(caret)
library(geojsonR)
library(gganimate)
library(gifski)
library(gridExtra)
library(kableExtra)
library(knitr)
library(lubridate)
library(mapview)
library(riem)
library(scales)
library(sf)
library(spdep)
library(tidycensus)
library(tidyverse)
library(tigris)
library(viridis)

# R options setup
options(scipen = 999)
options(tigris_class = "sf")
options(tigris_use_cache = T)
options(knitr.graphics.error = F)

# additional functions from PPA book
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")


# function shortcuts
g <- glimpse
m <- mapview
len <- length
st_c <- st_coordinates


# Aesthetic settings
colors <- c('#222222',
            '#eeeeee')

palette <- c("#676767",
             "#18b6c4",   # properties
             "#18c493",   # vacantLots
             "#f7c457",   # delinquencies
             "#f79c57",   # us bank liens
             "#f76957")   # sheriff sales

```


***

## Introduction



***


## Exploratory Analysis


```{r data: base geometry}

### A. Geographic boundaries

# set CRS to Pennsylvania South NAD83 in meters
phlcrs <- 'EPSG:32129'


# get geometry for tracts in Philadelphia
phltracts <- 
  tigris::tracts(state = 42, county = 101) %>%
  dplyr::select(GEOID, geometry) %>%
  st_transform(st_crs(phlcrs))


# get geometry for the county of Philadelphia
phlcounty <-
  phltracts %>%
  st_union()


# set up the fishnet bins as the most granular unit of analysis
fishnet <- phlcounty %>%
  st_make_grid(.,
               cellsize = 250, 
               square = TRUE) %>%
  .[phlcounty] %>%                     # clip to Philadelphia County boundary
  st_sf() %>%
  mutate(uniqueID = rownames(.))


# Create a fishnet grid boundary of Philadelphia County
phlFishnet <- fishnet %>%
  st_union()


### B. Demographic

# Get API key loaded locally
censusKey <-  read_file('../censusapikey.R')

# load census api key from local file
census_api_key(censusKey)

# set consulted year
year <- 2019

# get variables names and codes
acsVariableList <- load_variables(year, "acs5", cache = TRUE)

# set census variables to consult
censusVars <-
  c("B02001_001E", # Total population 
    "B11001_002E", # Total households
    "B02001_002E", # Total white population 
    "B19013_001E", # Median household income
    "B25003_001E") # Tenure of household

# Get the data
demographics <-
  get_acs(geography = "tract",
          variables = censusVars,
          year = year,
          state = 42,
          county = 101,
          geometry= T,
          output = "wide") %>%
  st_transform(st_crs(phlcrs)) %>%
  rename(totalPop = "B02001_001E",
         totalHHs = "B11001_002E",
         whitePop = "B02001_002E",
         medHHInc = "B19013_001E",
         tenureHH = "B25003_001E") %>%
  dplyr::select(-NAME, -starts_with("B"))%>%
  replace(is.na(.), 0) %>%
  mutate(pctWhite = ifelse(totalPop > 0, whitePop / totalPop, 0),
         pctTenur = ifelse(totalPop > 0, tenureHH / totalPop, 0)) %>%
  mutate(nhMajMin = ifelse(pctWhite > 0.5, 1, 0),
         nhIncome = ifelse(medHHInc > mean(.$medHHInc), 1, 0),
         nhTenure = ifelse(pctTenur > 0.5, 1, 0)) %>%
  dplyr::select(nhMajMin, nhIncome, nhTenure) %>%
  st_transform(st_crs(phlcrs))
  

# Create a mask for the majority minority census tracts fishnet cells
majminFishnet <- 
  demographics %>%
  filter(nhMajMin == 1) %>%
  st_union() %>%
  st_intersection(fishnet) %>%
  fishnet[.,] %>%
  st_union()


### C. Properties

# [source]('https://www.opendataphilly.org/dataset/opa-property-assessments')

# Load properties from 1999 until now from the Office of Property Assessment.
# properties <- read_csv('https://opendata-downloads.s3.amazonaws.com/opa_properties_public.csv')

# OR local load
properties <- readRDS('../data/local/properties.rds')

# Selected variables of interest
propertiesVars <-
  c("category_code",                  # KEEP ------------------------------------determines if it is VACANT LAND
    #"exterior_condition",             # KEEP ------------------------------------how the exterior appears based on observation.
    #"frontage",                       # MODEL
    #"location",                       # KEEP for JOIN check****
    "market_value",                   # KEEP ------------------------------------the certified market value of the property.
    #"off_street_open",                # UNNECESSARY
    "owner_1",                        # KEEP
    "owner_2",                        # KEEP
    "parcel_number",                  # KEEP to JOIN***
    #"parcel_shape",                   # MODEL
    #"registry_number",                # KEEP to JOIN
    "sale_date",                      # KEEP
    #"sale_price",                     # KEEP
    #"taxable_land",                   # MODEL
    #"topography",                     # MODEL
    "total_area",                     # KEEP
    #"unfinished",                     # UNNECESSARY
    "year_built",                     # KEEP
    "zoning",                         # KEEP
    "pin",                            # KEEP to JOIN ****
    "lat",                            # KEEP****
    "lng")                            # KEEP****


# Data wrangling
propertiesData <- properties %>%
  dplyr::select(propertiesVars) %>%
  filter(!is.na(lat), !is.na(lng)) %>%              
  st_as_sf(coords = c("lat","lng"), crs = 4326) %>%
  st_transform(st_crs(phlcrs)) %>%
  st_join(demographics)


# save locally
# saveRDS(properties, file = "properties.rds")

```


---

### Properties and Vacant Lots


```{r data: vacant land}

# [source]('https://www.opendataphilly.org/dataset/vacant-property-indicators')

# OpenDataPhilly Vacant Property Indicator - Lots
vacantLand <- read_csv('https://opendata.arcgis.com/datasets/19c35fb02d544a9bad0032b58268c9f9_0.csv')

# Select useful variables --- NO GEOMETRY
vacantLandVars <-
  c(#"ADDRESS",                        # KEEP for JOIN CHECK
    "BLDG_DESC",                      # KEEP ------------------------------------Building description from OPA
    "OPA_ID",                         # KEEP to JOIN****
    #"ZONINGBASEDISTRICT",             # KEEP***
    "LAND_RANK")                      # KEEP?

# Data wrangling
vacantLandData <- vacantLand %>%
  dplyr::select(vacantLandVars) %>%
  mutate(vacant = 'vacant')


# JOIN to propertiesData
# propertiesData <---> vacantLandData
vacantLandProps <- propertiesData %>%
  inner_join(vacantLandData, by=c('parcel_number'='OPA_ID'))

# save locally
# saveRDS(vacantLand, file = "vacantLand.rds")
saveRDS(vacantLandProps, file = "vacantLandProps.rds")

```



---


### Tax Delinquent Vacant Lots

```{r supply tax delinquencies}

# [source]('https://www.opendataphilly.org/dataset/property-tax-delinquencies')

# Real Estate Delinquencies from OpenData Philly:
delinquencies <- read_csv('https://phl.carto.com/api/v2/sql?q=SELECT+*+FROM+real_estate_tax_delinquencies&filename=real_estate_tax_delinquencies&format=csv&skipfields=cartodb_id,the_geom,the_geom_webmercator')

# Variables - Tax Delinquencies (53 vars)
delinquenciesVars <-
  c("opa_number",                     # KEEP to JOIN****
    #"street_address",                 # KEEP to JOIN check****
    #"owner",                          # UNNECESSARY
    #"co_owner",                       # UNNECESSARY
    "total_due",                      # KEEP ------------------------------------The total amount owner owes.
    "is_actionable",                  # KEEP?*** --------------------------------Action can be taken against the delinquent property.
    "num_years_owed",                 # KEEP ***for adverse possession reference
    "most_recent_year_owed",          # KEEP for reference
    "oldest_year_owed",               # KEEP for reference
    "most_recent_payment_date",       # KEEP for reference
    #"year_of_last_assessment",        # KEEP for reference
    "total_assessment",               # KEEP for reference
    #"building_code",                  # ???? ------------------------------------Building codes describe the building.
    #"general_building_description",   # ???? ------------------------------------General description of how the building is used.
    #"building_category",              # ???? ------------------------------------Type of Building Group (residential, commercial, etc.).
    "sheriff_sale",                   # KEEP*** ---------------------------------Property is in the Sheriff Sale Process (any stage).
    "liens_sold_1990s",               # KEEP*** ---------------------------------Property was included in 1997 Lien Sale.
    "liens_sold_2015")                # KEEP*** ---------------------------------Property is included in Recent Lien Sales.


# Data wrangling
delinquenciesData <- delinquencies %>%
  dplyr::select(delinquenciesVars) %>%
  mutate(opa_number = sprintf("%09d", opa_number))


# JOIN to properties
# propertiesData <---> delinquenciesData
delinquenciesProps <- propertiesData %>%
  inner_join(delinquenciesData, by = c('parcel_number'='opa_number')) %>%
  st_sf()

# save locally
# saveRDS(delinquencies, file = "delinquencies.rds")

# map vacant properties (points)
delinquenciesVacantProps <-
  left_join(vacantLandProps,
            delinquenciesProps %>% st_drop_geometry() %>% mutate(delinquentStatus = 1),
            by = "parcel_number") %>%
  mutate(delinquentStatus = ifelse(is.na(delinquentStatus) == F, 1, 0),
         delinquentType = ifelse(liens_sold_1990s == T, 2, delinquentStatus), # delinquent type: 0 not, 1 yes, 2 us bank
         delinquentType = replace_na(delinquentType, 0))


```



---


### Sheriff Sales


```{r data: transfers}

# [source]('https://www.opendataphilly.org/dataset/real-estate-transfers')

# OpenDataPhilly Endpoint of Property Assessment History:
# transfers <- read_csv('https://opendata-downloads.s3.amazonaws.com/rtt_summary.csv')
# transfers <- read_csv('https://opendata.arcgis.com/datasets/88e5bc291b834606bd49f6fd6dca226e_0.csv')

# OR load locally
transfers <- readRDS('../data/local/transfers.rds')

# OpenDataPhilly - Transfers (48 variables)
transfersVars <- 
  c("document_type",                # KEEP*** -----------------------------------refers to type of Real Estate Transaction
    "display_date",                 # KEEP***
    "street_address",               # KEEP for JOIN check
    "grantors",                     # KEEP --------------------------------------seller (on deeds), or borrower (on mortgages)
    "grantees",                     # KEEP --------------------------------------buyer, recipient, new owner, or lien holder
    "total_consideration",          # KEEP?** -----------------------------------good exchanged for the real estate (usually money)
    #"assessed_value",               # KEEP?** -----------------------------------assess value by OPA
    #"common_level_ratio",           # UNNECESSARY -------------------------------fair market : assessment values ratio
    #"fair_market_value",            # KEEP?** -----------------------------------assessment value by common level ratio
    #"receipt_date",                 # UNNECESSARY - display_date is used
    #"recording_date",               # UNNECESSARY - display_date is used
    #"document_date",                # UNNECESSARY - display_date is used
    "opa_account_num")              # KEEP****to join
    #"property_count")               # KEEP --------------------------------------Number of properties in document


# Data wrangling
transfersData <- transfers %>%
  dplyr::select(all_of(transfersVars)) %>%
  filter(!is.na(opa_account_num)) %>%
  filter(document_type %in% c('DEED',
                              'DEED LAND BANK',
                              'DEED MISCELLANEOUS',
                              'DEED MISCELLANEOUS TAXABLE',
                              'DEED OF CONDEMNATION',
                              'DEED SHERIFF'))

# JOIN to properties
# propertiesData <---> transfersData

# if a parcel_number has multiple deeds, I manipulated data to shift deed sheriff first
# so distinct() will grab deed sheriff and neglect others
transfersProps <- propertiesData %>%
  inner_join(transfersData, by= c('parcel_number'='opa_account_num'))%>% 
  mutate(document_type = factor(document_type,
                                labels(c("DEED SHERIFF" = 1,
                                         "DEED" = 2,
                                         "DEED MISCELLANEOUS" = 3,
                                         "DEED OF CONDEMNATION" = 4,
                                         "DEED LAND BANK" = 5,
                                         "DEED MISCELLANEOUS TAXABLE" = 6))))

  
# save locally
# saveRDS(transfers, file = "transfers.rds")

```





```{r data: sheriff sales, fig.width= 6, fig.height= 6}

# sheriff sales after 2021 from client
sheriffSales_21 <- read_csv('./data/sheriffSales_21.csv') %>%
  distinct(OPA, .keep_all=TRUE)                                                # remove duplicates

# Sheriff Sales before 2021 from real estate transfers data (completed)
sheriffSales_20 <-
  transfersProps %>%
  filter(document_type == "DEED SHERIFF") %>%
  arrange(parcel_number, document_type) %>%
  distinct(parcel_number, .keep_all = TRUE)


# join all sheriff sales info we have together
## sheriff sales data from transfers
sheriffProps <-
  left_join(
    delinquenciesVacantProps,
    st_drop_geometry(sheriffSales_20),
    by = "parcel_number") %>%
  distinct(parcel_number, .keep_all = T) %>%
  merge(., sheriffSales_21,
        by.x = "parcel_number", by.y = "OPA", all.x = T, no.dups = T) %>%
  mutate(pastSheriffSale = ifelse(document_type == "DEED SHERIFF", 1, 0)) %>%   #  past records only
  mutate(pastSheriffSale = replace_na(pastSheriffSale, 0)) %>%
  mutate(futureSheriffSale = ifelse(document_type == "DEED SHERIFF" | is.na(Status) == F | sheriff_sale == "Y", 1, 0)) %>% # future
  mutate(futureSheriffSale = replace_na(futureSheriffSale, 0)) %>%
  mutate(allSheriffSales = ifelse(pastSheriffSale == 1 | futureSheriffSale == 1, 1, 0),
         sheriffSaleYear = year(as.Date(display_date))) 


```


## Demand-side data

### L&I Building and Zoning Permits


```{r data: permits}

# [source]('https://www.opendataphilly.org/dataset/licenses-and-inspections-building-permits')

# Load PERMITS data from OpenDataPhilly's Carto API
# permits <- read_csv('https://phl.carto.com/api/v2/sql?q=SELECT+*,+ST_Y(the_geom)+AS+lat,+ST_X(the_geom)+AS+lng+FROM+permits&filename=permits&format=csv&skipfields=cartodb_id')

# OR local load
permits <- readRDS('../data/local/permits.rds')


# List the permit data set variables needed
permitsVars <-
  c('parcel_id_num',                 # KEEP for JOIN
    'permittype',                    # KEEP*****
    'permitdescription',             # KEEP for reference
    #'commercialorresidential',        # MAYBE
    'typeofwork',                    # KEEP****
    #'approvedscopeofwork',           # UNNECESSARY - detailed description
    'permitissuedate',               # KEEP***
    'status',                        # KEEP
    'applicanttype',                 # MAYBE
    #'contractor[...]',               # UNNECESSARY
    'opa_account_num',               # KEEP for JOIN****
    #'address',                       # KEEP for JOIN check
    #'unit_type',                     # UNNECESSARY
    'lng','lat'
    #'geometry'
    )

# EXAMINE VARIABLES
# permittype
permitTypes <- as.data.frame(table(permits$permittype))

# select permittyoe
permitCats <- 
  c('BP_NEWCNST',             #    4984   # "2007-01-02" to "2020-02-20"
    'BUILDING',               #    8521   # "2015-01-12" to NOW ("2022-01-29")
    'RESIDENTIAL BUILDING',   #   15909   # "2015-01-06" to NOW ("2022-01-29")
    'ZONING',                 #   11187   # "2015-12-02" to NOW ("2022-01-29")
    'ZP_ADMIN',               #     803
    'ZP_USE',                 #    9651   # "2007-01-02" to "2020-03-12"
    'ZP_ZON/USE',             #   12921   # "2007-01-02" to "2020-03-12"
    'ZP_ZONING'               #    7617   # "2007-01-02" to "2020-03-12"
    )

# Old categories
permitCatsA <- 
  c('BP_NEWCNST',             #    4984   # "2007-01-02" to "2020-02-20"
    'ZP_ADMIN',               #     803
    'ZP_USE',                 #    9651   # "2007-01-02" to "2020-03-12"
    'ZP_ZON/USE',             #   12921   # "2007-01-02" to "2020-03-12"
    'ZP_ZONING')              #    7617   # "2007-01-02" to "2020-03-12"

# New categories
permitCatsB <- 
  c('BUILDING',               #    8521   # "2015-01-12" to NOW ("2022-01-29")
    'RESIDENTIAL BUILDING',   #   15909   # "2015-01-06" to NOW ("2022-01-29")
    'ZONING')                 #   11187   # "2015-12-02" to NOW ("2022-01-29")


# Building categories
permitBuildingCats <- 
  c('BP_NEWCNST',             #    4984   # "2007-01-02" to "2020-02-20"
    'BUILDING',               #    8521   # "2015-01-12" to NOW ("2022-01-29")
    'RESIDENTIAL BUILDING')   #   15909   # "2015-01-06" to NOW ("2022-01-29")

# Zoning categories
permitZoningCats <- 
  c('ZP_ADMIN',               #     803
    'ZP_USE',                 #    9651   # "2007-01-02" to "2020-03-12"
    'ZP_ZON/USE',             #   12921   # "2007-01-02" to "2020-03-12"
    'ZP_ZONING',              #    7617   # "2007-01-02" to "2020-03-12"
    'ZONING')                 #   11187   # "2015-12-02" to NOW ("2022-01-29")


# Data wrangling ONE
# Filter data with PERMIT TYPES (building or zoning)
permitsData <- permits %>%
  dplyr::select(permitsVars) %>%
  filter(permittype %in% permitZoningCats) %>%
  filter(!is.na(lat), !is.na(lng)) %>% 
  st_as_sf(coords = c("lng","lat"), crs = phlcrs) 

# type of work across SELECTED categories (48):
permitTypesOfWork <- as.data.frame(table(permitsData$typeofwork)) 

# All categories
permitTypeOfWorkCats <- 
  c(#'ADD',                                              #   1665 ***OLD ***ZONING {New construction attached or added to existing}
    #'ADDITION AND/OR ALTERATION',                       #  17051 ***NEW ***BUILDING (6290)  ***RESIDENTIAL BUILDING (10761)
    #'CHANGE OF USE',                                    #   3046 ***NEW ***ZONING {General change of use}
    'COMBINED LOT LINE RELOCATION AND NEW DEVELOPMENT', #    313 ***NEW ***ZONING {Combine or redraw lots} ***SIGNIFICATIVE!
    'COMDEM',                                           #   1163 ***OLD ***ZONING {Complete demolition}
    'ENTIRE',                                           #   2910 ***OLD ***ZONING (-2890) ***BUILDING {Entire structure}
    'ENTSTR',                                           #   3638 ***OLD ***ZONING
    'FULL DEMOLITION',                                  #    644 ***NEW ***ZONING {Full demolition}
    'LOT LINE RELOCATION',                              #    317 ***NEW ***ZONING {Combine or redraw lots} ***SIGNIFICATIVE!
    'LOTLIN',                                           #   2527 ***OLD ***ZONING {Combine or redraw lots} ***SIGNIFICATIVE!
    'NEW CONSTRUCTION',                                 #<- 6292 ***NEW ***BUILDING (1812) ***RESIDENTIAL BUILDING (4480)
    'NEW CONSTRUCTION (SHELL ONLY)',                    #<-   19 ***NEW ***BUILDING
    #'NEW CONSTRUCTION (STAND ALONE)',                   #<-   48 ***NEW ***RESIDENTIAL BUILDING
    'NEW CONSTRUCTION, ADDITION, GFA CHANGE',           #<- 4986 ***NEW ***ZONING {New construction attached or added to existing}
    'NEWCON'#                                           #<- 5539 ***OLD ***ZONING (-3) ***BUILDING {New construction of structure}
    #'PARTCH',                                           #   3317 ***OLD ***ZONING {Change in use}
    #'SFADD',                                            #   3149 ***OLD ***ZONING {Add square footage to existing}
    )
    
# Filter data with PERMIT TYPES (building or zoning)
permitsZoningData <- permitsData %>%
  filter(typeofwork %in% permitTypeOfWorkCats)


# JOIN to properties
# permitsZoningData <---> propertiesData
permitsProps <- propertiesData %>%
  inner_join(st_drop_geometry(permitsZoningData),  by = c('parcel_number'='opa_account_num'))



```



---


## Development Activity






```{r fig.width= 8, fig.height=4}

allPermitsProps <-
  left_join(delinquenciesVacantProps,
            st_drop_geometry(permitsProps),
            by=c("parcel_number")) %>%
  distinct(parcel_number, .keep_all = TRUE)


tabPermitByYr <-
  allPermitsProps %>%
  filter(is.na(permittype) == F) %>%
  group_by(is.na(permittype), year(permitissuedate)) %>%
  summarise(countPermit = n()) %>%
  st_drop_geometry() 


tabPermitAllByYr <- 
  permitsProps %>%
  group_by(is.na(permittype), year(permitissuedate)) %>%
  summarise(countPermit = n()) %>%
  st_drop_geometry() 


tabPermitBothByYr <-
  merge(tabPermitByYr, tabPermitAllByYr, by="year(permitissuedate)") %>%
  dplyr::select(1,3,5) %>%
  rename("year" = 1,
         "permitOnVacant" = 2,
         "permitPhilly" = 3) %>%
  gather(key="variable", value="value", -year)


```


## Feature Engineering


### Permits 

```{r permits base}

# permits by quarter by fishnet cell
permitsNet <- permitsProps %>%
  st_join(fishnet, left=F) %>%
  mutate(year = year(permitissuedate),
         month = month(permitissuedate),
         quarter = case_when(
           month %in% c(1, 2, 3) ~ 1,
           month %in% c(4, 5, 6) ~ 2,
           month %in% c(7, 8, 9) ~ 3,
           month %in% c(10, 11, 12) ~ 4)) %>%
  mutate(period = paste(year, quarter, sep='-')) %>%
  group_by(period, uniqueID) %>%
  summarize(permitCount = sum(n())) %>%
  st_drop_geometry()

# random cross validation ID (out of 100) for later use
# mutate(uniqueID = rownames(.)) %>%
# mutate(cvID = sample(round(nrow(fishnet)/24), size=nrow(fishnet), replace=T)) 

# create empty panel with all possible time/space combinations by fishnet grids
basePanel <- 
  expand.grid(period = unique(permitsNet$period), 
              uniqueID = unique(fishnet$uniqueID))


# join panel to fishnet geometry
permitsPanel <- 
  permitsNet %>%
  right_join(basePanel) %>% 
  replace(is.na(.), 0) %>%
  left_join(fishnet) %>%
  st_sf()


# save locally
# saveRDS(permits, file = "permits.rds")

```


```{r permit engineering}


# classify zoning permits by five-year periods
permitsYears <- 
  permitsPanel %>%
  mutate(year = as.integer(substr(period, 1,4))) %>%
  filter(year != '2022') %>%
  dplyr::select(-period) %>%
  group_by(year, uniqueID) %>%
  summarize(permitCount = sum(permitCount))


# Select the four-year interval (16, 17, 18 and 19)
permitsNetYears <- 
  st_drop_geometry(permitsYears) %>%
  filter(year < 2020,
         year > 2015)


# classify zoning permits by five-year periods
permitsLustrums <- 
  permitsYears %>%
  mutate(lustrum = case_when(
           year %in% seq(2007, 2011) ~ '2007-2011',
           year %in% seq(2012, 2016) ~ '2012-2016',
           year %in% seq(2017, 2021) ~ '2017-2021'
         )) %>%
  dplyr::select(-year) %>%
  group_by(lustrum, uniqueID) %>%
  summarize(permitCount = sum(permitCount))

# save locally
# saveRDS(permitsYears, "permitYears.Rds")


```





### A. Sales over the last three years

```{r sales}

# variables to keep
salesVars <- 
  c("geometry",
    "document_type",
    "display_date",
    "grantors",
    "grantees",
    "total_consideration",
    "year",
    "period",
    "total_area")


# if a parcel_number has multiple sales, we only get the latest one
salesProps <-
  transfersProps %>%
  drop_na(total_area) %>% # drop sales that have NA for their area
  mutate(document_type = as.character(document_type)) %>%  # to ignore Gillian deed factor levels
  group_by(parcel_number) %>%
  slice(which.max(display_date)) %>% # to get only the latest one
  mutate(year = year(display_date),
         month = month(display_date),
         quarter = case_when(
           month %in% c(1, 2, 3) ~ 1,
           month %in% c(4, 5, 6) ~ 2,
           month %in% c(7, 8, 9) ~ 3,
           month %in% c(10, 11, 12) ~ 4)) %>%
  mutate(period = paste(year, quarter, sep='-')) %>%
  st_sf()


# select and transform features + filter by period + JOIN into fishnet
salesPropsNet <- 
  salesProps %>%
  dplyr::select(all_of(salesVars)) %>%
  filter(year > 2015,
         year < 2020,
         total_consideration < 4e8,
         total_consideration > 1e4) %>%  # filter outliers and most "symbolic" transfers
  mutate(sqftPrice = ifelse(total_area > 0, total_consideration/total_area, 0)) %>% # calculate price by sqft
  st_join(fishnet, left = F)


# create empty panel with all possible time/space combinations by fishnet grids YEARS
yearPanel <- 
  expand.grid(year = unique(salesPropsNet$year), 
              uniqueID = unique(fishnet$uniqueID))



# Get sales count by fishnet from 2016 to 2019
salesNetYears <- 
  salesPropsNet %>%
  group_by(year, uniqueID) %>%
  summarize(salesCount = sum(n())) %>%
  st_drop_geometry() %>%
  right_join(yearPanel) %>%
  replace(is.na(.), 0)

# Get sqft price paid by fishnet from 2016 to 2019
sqftNetYears <-
  salesPropsNet %>%
  group_by(year, uniqueID) %>%
  summarize(meanSqftPrice = mean(sqftPrice)) %>%
  st_drop_geometry()  %>%
  right_join(yearPanel) %>%
  replace(is.na(.), 0)

# Get total price paid by fishnet from 2016 to 2019
priceNetYears <-
  salesPropsNet %>%
  group_by(year, uniqueID) %>%
  summarize(meanPrice = mean(total_consideration)) %>%
  st_drop_geometry()  %>%
  right_join(yearPanel) %>%
  replace(is.na(.), 0)

# Get change in number of individual owners by fishnet from 2016 to 2019
individualOwnersNetYears <- 
  salesPropsNet %>%
  group_by(year, uniqueID) %>%
  summarize(numIndGrantors = n_distinct(grantors),
            numIndGrantees = n_distinct(grantees)) %>%
  mutate(indOwnerDelta = abs(numIndGrantors - numIndGrantees)) %>%
  dplyr::select(-numIndGrantors, -numIndGrantees) %>%
  st_drop_geometry()  %>%
  right_join(yearPanel) %>%
  replace(is.na(.), 0)





m(head(salesPropsNet, 100))





```



### B. Amount of delinquent debt

```{r debt}

# All delinquent properties
debtNetYears <- delinquenciesProps %>%
  filter(oldest_year_owed < 2020) %>% # take out the ones whose debt started after 2018
  rename('year' = most_recent_year_owed) %>%
  st_join(., fishnet) %>%
  group_by(year, uniqueID) %>%
  summarize(meanDebt = mean(total_due)) %>%
  filter(year > 2015) %>%
  st_drop_geometry()  %>%
  right_join(yearPanel) %>%
  replace(is.na(.), 0)

```

### C. Average Rent

```{r rent}

rentTracts <-
  get_acs(geography = "tract",
          variables = c("B25064_001"),
          year = 2018,
          state = "PA",
          geometry = TRUE,
          county=c("Philadelphia"),
          output = "wide") %>%
  rename('medianRent' = B25064_001E) %>%
  dplyr::select(medianRent, GEOID) %>%
  st_transform(st_crs(phlcrs))


m(rentTracts, zcol = 'medianRent')


fishnetCentroids <-
  fishnet %>%
  st_centroid()


```


### D. Demographic Change (Gillian)

Total population, pct of white population, income by household.

```{r demographic change}

```




### F. Number of vacant lots

```{r vacancy}

# get count of vacant lots and cumulative vacant area
# ***THIS DOES NOT HAVE YEAR DIFFERENCE
vacantNet <- 
  vacantLandProps %>%
  drop_na(total_area) %>% # drop sales that have NA for their area
  st_sf() %>%
  st_join(., fishnet) %>%
  group_by(uniqueID) %>%
  summarize(cumulativeVacantArea = sum(total_area),
            totalVacantLots = sum(n())) %>%
  st_drop_geometry() %>%
  right_join(yearPanel) %>%
  replace(is.na(.), 0)

```

### H. Sheriff Sales

```{r sheriff sales}

```


### E. Exposure to amenities

```{r exposure}


```

#### E1. Schools
```{r exposure schools}

```

#### E2. Parks
```{r exposure parks}

```

#### E3. Transit
```{r exposure transit}

```

#### E4. Jobs
```{r exposure jobs}

```

#### E5. Crime
```{r exposure crime}

```

#### E5. Commerce
```{r commerce}

```





###. Complete feature set

```{r join features}

completeNet <- 
  permitsNetYears %>%
  left_join(salesNetYears) %>%
  left_join(priceNetYears) %>%
  left_join(sqftNetYears) %>%
  left_join(debtNetYears) %>%
  left_join(individualOwnersNetYears) %>%
  left_join(vacantNet)

```




```{r}


gilliansFeatures <- readRDS("local/featureNet.rds")

gilliansFeatures


```






```{r}

animation <- 
  ggplot() +
  geom_sf(data = salesPanel, aes(fill = permitCount), color = NA) +
  scale_fill_viridis(option = "magma",
                     limits = c(0, 20),
                     #trans = "log1p",
                     breaks = c(0, 5, 10, 20),
                     direction = 1) +
  labs(title = "Sales in Philadelphia",
       fill = "",
       caption = "Source: OpenDataPhilly") +
  guides(fill= guide_colorbar(barheight = 9, title.position = "bottom", label.position = "right")) +
  geom_text(data = salesPanel,
            aes(label = substr(period, 1, 4)),
            size = 8, x = -Inf, y = Inf, vjust = 1.5, hjust = -.1, colour = "black") +
    transition_manual(period) +
    mapTheme() +
    theme(legend.position = c(0.85, 0.2),
          legend.background = element_blank(),
          panel.border = element_blank(),
          panel.background = element_rect(fill = colors[2]),
          panel.grid = element_blank()
          )

animation

animate(animation, duration=20, renderer = gifski_renderer())

anim_save("sales", animation, duration=20, renderer = gifski_renderer())


```






